### 2018-06-13 highlights.tex
submission

\documentclass[11pt,a4paper]{article}        % possible options include font size ('10pt', '11pt' and '12pt'), paper size ('a4paper', 'letterpaper', 'a5paper', 'legalpaper', 'executivepaper' and 'landscape') and font family ('sans' and 'roman')

\usepackage[document]{ragged2e}
\usepackage[utf8]{inputenc}              
\begin{document}

\section*{Highlights}

\begin{itemize}
	\item We investigate the sensitivity of geosimulation models to initial spatial conditions.
	\item We generate stylized population density grids.
	\item We assess the sensitivity of two classical agent-based models, Schelling's model and Sugarscape's model.
	\item Strong quantitative and qualitative effects of spatial initial conditions are shown for these models.
\end{itemize}

\end{document}

%% end of file `template.tex'.